---
title: "Visualization Lab 6"
author: "Roshni Sundaramurthy (rossu809) & Brian Masinde (brima748)"
date: "04 October 2018"
output:
  html_document:
    df_print: paged
    self_contained: no
  #html_notebook:
  #  theme: journal
  #pdf_document: default
fontsize: 11pt
bibliography: references.bib
---

## Assignment 1. Network Visualization of terrorist connections

```{r libs, message=FALSE, warning=FALSE, paged.print=FALSE}
library(ggraph)
library(igraph)
library(visNetwork)
library(htmlwidgets)
library(seriation)
library(plotly)
```


```{r data}
#_______________________LINKS___________________________________________
# connections between individuals (first two columns)
# strength of ties linking them 1:4
trainData <- read.table("trainData.dat")

colnames(trainData) <- c("from","to","weight")

trainData$width <- trainData$weight

# convert cols to factors
#trainData$from <- as.factor(trainData$from)

#trainData$to <-as.factor(trainData$to)

#trainData$tie_strength <- as.factor(trainData$tie_strength)

#__________________NODES________________________________________________
# 
trainMeta <- read.table("trainMeta.dat")

# rename columns
colnames(trainMeta) <- c("person","bomb_group")

# create an id column
trainMeta$id <- 1:nrow(trainMeta)

# convert part (participation) to factor
trainMeta$bomb_group <- as.factor(trainMeta$bomb_group)

# label column
trainMeta$label = trainMeta$person
```

### 1
```{r message=FALSE, warning=FALSE, paged.print=FALSE, fig.width = 9 , fig.height =9}
#_____________link order [from and to]________________________________
trainData <- trainData[order(trainData$from, trainData$to),]

#______________________NODE___________________________________________
node <- data.frame(
  id = trainMeta$id,
  label = trainMeta$label,
  group = trainMeta$bomb_group
  #value =  strength(graph = g)
)

#____________________EDGES_____________________________________________

edge <- data.frame(
  from = trainData$from,
  to = trainData$to,
  weight = trainData$weight,
  width = trainData$width
)

#________________________________GRAPH_______________________________
# graph object use graph_from_data_frame(), id needs to be first

g <- graph_from_data_frame(
  d = edge,
  vertices = node
)

#___________________________add value to node_____________________

node$value <- strength(graph = g)

graph_obj <- visNetwork(node, edge, width = "100%") %>%
  visPhysics(
    solver = "repulsion"
  ) %>%
  visLegend() %>% visOptions(highlightNearest = TRUE, nodesIdSelection = TRUE) 

graph_obj
```

From the network, we can observe three main clusters. The main cluster has **Jamal Zougam**, as center of the cluster. In particular, observing the size of the node that represents **Jamal Zougam**, he has the highest number of connections. Also a person of interest would be **Mohamed Chaoui**, as he has the second most largest connections. 

The second cluster is comprised of a group that do not have any direct relationship to the bombing group. This cluster is comprised of ** Emilio Liamo**, ** Jose Emilio Suarez**, ** El Gitalino**, **Antonio Toro**, **Ivan Granados**, **Raul Gonzalez Perez**.

There is a group of people who have no connection at all to anyone. Such people are: **Faisal Alluch**, **Sanel Sjekirika**,  and **Abdelhalak Bentasser**

### 2.

```{r qn2, fig.height = 9, fig.width =9}
#_________________________Path length one or two_________________________________

graph_obj %>%
  visOptions(highlightNearest = list(enabled = TRUE, degree = 2), nodesIdSelection = TRUE)
```

Jamal Zougam and Mohamed Chaoui appear to have the same links with path lenght one or two. However, since the former has the largest connection (size of the representative node), he has best opportunity to spread information. From wikepedia, Jamal Zougam had a telephone shop and it was established that the phones from his store were used in the train bombings. It is a possiblity that he also distributed phones in the network to facilitate communication.

### 3.

```{r qn3}
#___________________________________Clusters_________________________________________

# make a copy of nodes, no need to overwrite it
nodes1<-node

# a graph object
net <- graph_from_data_frame(d=edge, vertices=node, directed=F)

# cluster
ceb <- cluster_edge_betweenness(net)

# a group variable from edge betweenness score (number of shortest paths through it)

nodes1$group=ceb$membership

visNetwork(nodes1,edge)%>%
  visIgraphLayout()
```

The clustering method by calculating edge betweenness score identifies all the clusters we were able to identify. **Jamal Zougam** is a hub in this graph too. We note that there are more clusters than we would have been able to identify observation. Due to the coloring in this output there are 7 clusters.

### 4.

```{r}
#_______ Permutation byHC____________________________________________

# adjacency representation


netm <- get.adjacency(net, attr="weight", sparse=F)

# colnames
colnames(netm) <- V(net)$label

#rownames
rownames(netm) <- V(net)$label

# calculate distances
rowdist<-dist(netm)

# get order

#set seed

set.seed(111)
order1<-seriate(rowdist, "HC")

ord1<-get_order(order1)

reordmatr<-netm[ord1,ord1]

plot_ly(z=~reordmatr, x=~colnames(reordmatr), 
        y=~rownames(reordmatr), type="heatmap")
```

The most pronounced cluster would be at the bottom left coner of the heatmap. The persons in this cluster have ties to
Al Qaeda, and to Osama Bin Laden. Example of such people are: Suresh Kumar, Vinay Kholy, and Basel Ghayoun among others. This cluster was identified in Step 3 as the nodes colored red. In step 1 this cluster is not easily identified because we have to observe the thickness of the edges and the links.

